\begingroup
\graphicspath{{../prob_1/Figures/}}

% Problem 1: Feature extraction and MLP benchmark
\section{Feature Extraction and MLP Benchmark}

\subsection{Dataset and Preprocessing}
The Reduced MNIST dataset was downloaded through \texttt{kagglehub} and already provided in labeled training and testing folders, one subdirectory per digit class. In our implementation, each image was loaded from disk, converted again to grayscale to ensure a single-channel input, and normalized to the range $[0,1]$ by dividing pixel intensities by $255$.

The label of each sample was taken directly from the folder name. No relabeling or manual reshuffling was introduced, which keeps the preprocessing aligned with the assignment dataset structure.

\subsection{Feature Construction}
The assignment required three feature representations: DCT, PCA, and autoencoder features. We implemented each feature extractor separately so that the same classifier could be evaluated on all representations under a fair comparison.

\subsubsection{Discrete Cosine Transform Features}
For the DCT representation, each grayscale image was transformed using a two-dimensional discrete cosine transform with \texttt{scipy.fft.dctn}. We used the orthonormal DCT-II variant and retained the top-left $15\times15$ low-frequency block, which captures the most informative spatial structure while reducing the original $28\times28$ image to a compact 225-dimensional vector.

\subsubsection{Principal Component Analysis Features}
For PCA, each image was first flattened into a 784-dimensional vector and then projected onto the principal subspace using \texttt{sklearn.decomposition.PCA}. We configured PCA with \texttt{n\_components=0.95}, which preserves 95\% of the variance in the training data.

\subsubsection{Autoencoder Features}
We implemented it as part of the same feature pipeline so that its learned latent space could be compared directly with DCT and PCA.

The autoencoder is implemented as a multilayer perceptron regressor with symmetric hidden layers \texttt{(128, 64, 32, 64, 128)}. It is trained to reconstruct the input image using ReLU activations and early stopping. The bottleneck is the middle hidden layer with 32 units. After training, we pass the data forward layer by layer and extract the bottleneck activations as the learned feature vector for each image.

\subsection{Classifier Design}
After extracting features, we trained a multilayer perceptron classifier on each feature set. The classifier uses ReLU activations, Adam optimization, early stopping, and a maximum of 120 iterations.

We evaluated three hidden-layer configurations:
\begin{itemize}
  \item one hidden layer: \texttt{(512,)}
  \item three hidden layers: \texttt{(512, 256, 128)}
  \item four hidden layers: \texttt{(512, 256, 128, 64)}
\end{itemize}

These configurations were chosen to study the effect of model depth on classification performance while keeping the feature extractor fixed.

\subsection{Integration with the Assignment 1 Benchmark}
The autoencoder implementation was also reused later in Problem 2 because the assignment table required us to rerun the Assignment 1 Part 2 benchmark using this feature.

\subsection{Summary}
In summary, Problem 1 focuses on the feature-extraction stage of the experiment. The three required features were implemented exactly as separate pipelines, and the same MLP classifier was applied to all of them using three hidden-layer configurations. The detailed accuracy and timing results are documented in the results table of Problem 2.

\endgroup